Illusory Finger Stretching and Somatosensory Responses

### 2024-12-02 manuscript.qmd
Update manuscript.qmd
Effect sizes updated

# install R packages
packagelist <- c('tidyverse','ggpubr','ez','ggplot2','jpeg','grid','plotrix','lavaan', 'cowplot', 'ggpubr','rstatix','reticulate','moments','tiff','rmarkdown','tinytex','osfr','FSA')  
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

@@

@@

@@

@@

@@

Resizing illusions, delivered using augmented reality, resize a body part through either stretching or shrinking manipulations. These resizing illusions have been investigated in visuotactile, visual-only, and visuo-auditory presentations. However, the neural underpinnings of these resizing illusions remain undefined. This study sought to understand the neural mechanisms behind these illusions (n = 46), by using somatosensory steady state evoked potentials in addition to subjective self-report questionnaires, to enhance knowledge of what drives the subjective embodiment during resizing illusions. Since these Illusions have been shown to provide analgesic effects for individuals with chronic pain conditions, this study also aimed to provide an empirical basis for future investigations in chronic pain samples undergoing resizing illusions. Confirmatory results demonstrated significant differences in subjective experience between non-illusion and multisensory illusion conditions, whilst EEG data measuring somatosensory response across electrodes of interest (F1 & FC1) to 26Hz stimulation to the resized digit showed no significant effects of condition. Exploratory non-parametric SSEP analyses showed a significant effect of condition, with reduced amplitudes in illusory conditions compared to non-illusion conditions, however no significant differences arose in exploratory post hoc tests. These findings demonstrate no clear effect of resizing illusions on SSEP amplitudes for participants without chronic pain, but exploratory findings could hint to a potential sharpening of neural representations as a result of illusory stretching, thereby providing a basis for investigations of comparable subjective and steady state illusion responses in a chronic pain population, who are thought to have more diffuse neural representations of their affected body parts.

@@

The augmented reality system used to deliver these resizing illusions presents real-time video capture of the hand, from the same position and perspective as if the hand were being viewed directly [@preston2011a]. This allows the experimenter to deliver tactile manipulations, such as gently pulling or pushing the hand/finger, whilst the participant views their hand/finger either stretching or shrinking in the augmented image. Newport, Pearce and Preston (2010) found strong embodiment using a synchronous multisensory visuotactile illusion, which was replicated in our pilot data using the same experimental set up as the current study. The pilot data showed, although not significant, a numerically greater illusory experience during synchronous visuotactile manipulations compared to asynchronous (mismatching visuotactile manipulation) control conditions (Appendix A) for illusory finger resizing. When comparing multisensory visuotactile resizing illusions to unimodal visual resizing illusions, our recent work [@hansford2023a] shows that multisensory illusions elicit significantly greater illusory experience compared to non-illusion and unimodal visual illusion conditions in healthy participants. We also showed, in exploratory analysis, that a subset of participants who experienced an illusion in the unimodal visual condition reported a stronger illusory experience in this condition than in an incongruent (mismatching visual and tactile inputs) control condition. This subset analysis, however, was of a small sample size, and was selected based on one of the measures analysed thus should be taken with caution, meaning further replication of the findings are needed. Furthermore, we have demonstrated that a visuo-auditory presentation of the finger resizing illusion, using non-naturalistic auditory input, provides a stronger illusory experience than a visual only presentation, but this does not surpass the illusion strength given by a visuo-tactile illusion [@hansford2023b].

@@

@@

Overall, based on the power analyses in section 2.5, a total sample size of 46 participants were tested. This sample size adheres to the higher end of sample size estimates (Hypothesis 2 (2.5.2) showing 46 participants needed for post hoc tests 2a – 2c).

@@

Ethical approval for this research was gained from the Department of Psychology, University of York (ethics application code 950), in line with the Declaration of Helsinki. All participants gave informed consent prior to the start of any experimental set up, and participants were instructed that they could withdraw their participation at any time during or after completion of the experiment. 46 participants were tested, with the participant’s manipulated finger being randomly split between use of either the index or middle finger. However, 2 participant’s data needed removal (>50% of electrodes requiring removal), and therefore 2 additional participants were tested to account for this missing data, both using the index finger as the manipulated digit, resulting in a final sample size of 46 participants (37 Female, 8 Male, 1 Prefer not say; Mean age = 20.3 years, age range = 18.3 – 32.7 years; 32 White, 11 Asian or Asian British, 3 Mixed or Multiple Ethnic Groups; Sample population = students at the University of York). 23 participants were tested using their index finger, the other half using their middle finger.

Sample inclusion / exclusion criteria:

@@

-	Less than 100% of the experiment completed by a participant, more than 50% of electrodes for a single participant requiring removal from EEG data, or if both electrodes F1 and FC1 (electrodes of interest) required removal. More information about data removal can be found in section 2.4.1 Preprocessing Steps.

@@

The experimenter was seated opposite the participant, the other side of the augmented reality machine and touched the digit during MS and NIT conditions by holding onto the distal interphalangeal joint and gently touching (NIT) or pulling (MS) the finger whilst the participant kept their hand in place. Conditions were delivered across 4 blocks, with each block consisting of 24 trials of the same experimental condition, totalling 96 trials over all 4 blocks. The ordering of the blocks was randomised for each participant to prevent ordering effects. The experiment was programmed in, and the conditions randomised using MATLAB R2017a and the experimenter was informed of whether to pull the finger or to touch the finger via an indicative box displayed on the screen out of the participant’s view. If the box was blue, this indicated a need to pull the finger, if it was white it indicated a need to touch the finger, if there was no box displayed then this indicated no tactile manipulation from the experimenter. The researcher used a button press to trigger the start of the manipulation, and started pulling the finger, when needed, synchronously within the 2.4 second manipulation phase. If the experimenter were to forget to pull the finger on a multisensory condition, or mistakenly pulled the finger in a control trial, then this would noted during the experiment, and that trial would be removed from analysis. Fortunately, no trials needed removal due to experimenter error. Vibrations were delivered to the participant’s finger in all conditions using a miniature electromagnetic solenoid stimulator (Dancer Design Tactor; diameter 1.8mm) emitting vibrations produced by sending amplified 26Hz sine wave sound files, with stimulus intensity controlled by an amplifier (Dancer Design TactAmp). The tactor was driven at 50% of the maximum (i.e. a peak input voltage of 3V) using a 26Hz sine-wave, and delivered a peak force of 0.18N. The electromagnetic solenoid stimulator was attached to the participant’s finger that was outstretched to receive the manipulations, between the knuckle and the first finger joint, using clear medical tape and gave continuous stimulation for the duration of each trial. Participants were encouraged to take a break between each of the blocks to stretch their hand. EEG was recorded throughout as a continuous recording with conditions denoted by numbered 8-bit digital at the start of each trial (USB-TTL Module, Black Box Toolkit Ltd.).

Finally, at the end of each block, the participant was asked to complete the subjective illusory experience questionnaire regarding a condition presented in a given block using a Samsung Galaxy Tab A6 tablet via a questionnaire on Qualtrics (Qualtrics, Provo, UT). The questionnaire consisted of six questions relating to the trials the participant had just experienced. Two statements related to illusory experience: “It felt like my finger was really stretching” / “It felt like the finger I saw was part of my body”, two related to disownership: “It felt like the finger I saw no longer belonged to me” / “It felt like the finger I saw was no longer part of my body”, and two were control questions: “It felt as if my finger had disappeared” / “It felt as if I might have had an extra finger” (all questions were directed towards the participants manipulated finger). Control questions were included to create an index for the illusion and disownership questions (more detail can be found in section 2.4.1 - Preprocessing steps), whilst disownership questions were included to assess if the potential experience from the illusions resulted from a disownership of the body part, or from subjective embodiment of the body part [@mccabe2011a]. A visual analogue scale from 0 – 100 was used for each statement, with 0 being strongly disagree, 50 being neutral and 100 being strongly agree.

@@

## Analysis Pipeline

### Preprocessing Steps

@@

Regarding questionnaire data, scores for both illusion experience questions were combined to give median scores, along with both disownership questions and both control questions, resulting in 3 median scores per condition per participant. The median control scores were used to create an index of the illusion and disownership scores by subtracting the median control score from the median illusion and median disownership scores, in line with previous research doing similarly [@matsumiya2021a; @kilteni2017a; @kalckert2012a]. The normalised (control indexed) data were used for analyses, with a new scale from -100 to +100 with 100 indicating strongly agree, 50 indicating a neutral opinion, and scores below 0 indicating strongly disagree with the statements on the questionnaire. 50 is maintained as a neutral opinion so that the normalised data still adhered to the thresholds that the participants were presented with during the experiment.

## Planned Analyses

*(1 – Positive Control) There will be a greater illusory experience, measured via a subjective illusory experience questionnaire, in the (1a) MS condition compared to the NI condition and in the (1b) MS condition compared to the NIT condition.*

The subjective illusory experience questionnaire was  used as a positive control for the current study. Previous research has shown significantly greater illusion strength for MS conditions compared to non-illusion conditions (Carey et al., 2019; Hansford et al., 2023a), which we attempted to replicate. Questionnaire data were be analysed using R (R Core Team, 2021). A one-way ANOVA would be run to compare the dependent variable of normalised (control indexed) illusion score from each independent condition. Given significant findings, post-hoc tests would be run, with Bonferroni correction for 2 comparisons (MS Vs NI, MS Vs NIT) at an initial alpha of 0.05.

Interpretations for hypothesis 1 can be found in the design table (Table 1).

*(2) There will be a significant difference in somatosensory SSEP response across the electrodes of interest (F1 & FC1). Subsequent hypotheses are that there will be a significant difference in SSEP response in (2a) the MS condition compared to the NI condition, and (2b) the UV condition compared to the NI condition, but (2c) that there will be no significant difference when comparing the NIT condition to the NI condition. *

As mentioned in the EEG pre-processing steps in section 2.4.1, analysis of EEG data involved taking a Fourier transform for each waveform averaged across the electrodes of interest, to obtain the amplitude for each trial at the vibration frequency (26Hz). These amplitudes would then be averaged across trials to give overall results for each participant, before running a repeated measures one way ANOVA comparing somatosensory SSEP response from each experimental condition. The dependent variable would be SSEP amplitude in µV, whilst the independent variable would be the different manipulations given in each comparison condition. Given significant findings in the ANOVA, post hoc comparisons would be conducted at a new alpha of .016 (corrected for 3 comparisons (MS Vs NI, UV Vs NI, NIT Vs NI)). Based on the pilot data in Appendix A, we expected to see activation most pronounced over mid-frontal distributions, covering F1 and FC1 electrodes and therefore these electrodes were selected as the electrodes of interest.

Interpretations for hypothesis 2 can be found in the design table (Table 1).

@@

The stage 1 report for this study, with details regarding planned analyses and pilot data, can be seen at the following OSF page: https://osf.io/pfksu/ . All data and code to reproduce the analyses can be found at the following OSF page:  https://osf.io/yhz6j/. A design planner encompassing research questions, hypotheses, sampling and analysis plans and their resulting interpretations and observed outcomes can be seen in Figure 3.

![Design Planner](Figures/designplanner.pdf){#fig-designplanner}

@@

Positive control analyses of the subjective illusion data can be seen in @fig-illusion. A one-way ANOVA found a significant overall effect of condition with a large effect size (*F*(`r round(res.aov1$ANOVA$DFn[1])`,`r round(res.aov1$ANOVA$DFd[1])`), *p.adj* = `r format.pval(res.aov1$ANOVA$p[1], digits = 1, eps = 0.001, nsmall = 3)`, $ηp^2$ = `r res.aov1$ANOVA$pes[1]`). Post hoc t tests with Bonferroni corrections found significantly greater combined illusion score in the MS condition (M = `r round(illusionmsds$mean[3],digits = 2)`, SD = `r round(illusionmsds$sd[3],digits = 2)`) compared to the Non-Illusion (NI; M = `r round(illusionmsds$mean[1],digits = 2)`, SD = `r round(illusionmsds$sd[1],digits = 2)`, *t* = `r round(pwc1$statistic[2], digits = 2)`, *p.adj* = `r format.pval(pwc1$p.adj[2], digits = 1, eps = 0.001, nsmall = 3)`, Cohen's d = `r round(pwc1$estimate[2])`) and Non Illusion Tactile (NIT; M = `r round(illusionmsds$mean[2],digits = 2)`, SD = `r round(illusionmsds$sd[2],digits = 2)`, *t* = `r round(pwc1$statistic[4], digit = 2)`, *p.adj* = `r format.pval(pwc1$p.adj[4], digits = 1, eps = 0.001, nsmall = 3)`, Cohen's d = `r round(pwc1$estimate[4])`) conditions, thereby supporting hypotheses 1, 1a, and 1b and fulfilling the positive control checks.

@@

Analyses of SSEP data can be seen in @fig-eeg1. The left panel confirms the presence of a clear steady-state signal at 26Hz, which was strongest over the fronto-central electrodes. A one-way ANOVA found no significant effect of condition with a small effect size (*F*(`r round(res.aov4$ANOVA$DFn[1])`,`r round(res.aov4$ANOVA$DFd[1])`), *p* = `r format.pval(res.aov4$ANOVA$p[1], digits = 1, eps = 0.001, nsmall = 3)`, $ηp^2$ = `r res.aov4$ANOVA$pes[1]`), opposing Hypothesis 2. Post hoc t tests with Bonferroni corrections found no significant differences between SSEP amplitude when comparing the NI condition (M = `r round(eegdatamsds$mean[1],digits = 2)`, SD = `r round(eegdatamsds$sd[1],digits = 2)`) to the MS condition (M = `r round(eegdatamsds$mean[3],digits = 2)`, SD = `r round(eegdatamsds$sd[3],digits = 2)`, *t* = `r round(pwc4$statistic[2], digits = 2)`, *p.adj* = `r format.pval(pwc4$p.adj[2], digits = 1, eps = 0.001, nsmall = 3)`, Cohen's d = `r round(pwc4$estimate[2],digits = 2)`), or UV condition (M = `r round(eegdatamsds$mean[4],digits = 2)`, SD = `r round(eegdatamsds$sd[4],digits = 2)`, *t* = `r round(pwc4$statistic[3], digits = 2)`, *p.adj* = `r format.pval(pwc4$p.adj[3], digits = 1, eps = 0.001, nsmall = 3)`, Cohen's d = `r round(pwc4$estimate[3],digits = 2)`), meaning Hypotheses 2a and 2b were unsupported. There was no significant difference found when comparing the NI condition to the NIT condition (M = `r round(eegdatamsds$mean[2],digits = 2)`, SD = `r round(eegdatamsds$sd[2],digits = 2)`, *t* = `r round(pwc4$statistic[1], digits = 2)`, *p.adj* = `r format.pval(pwc4$p.adj[1], digits = 1, eps = 0.001, nsmall = 3)`, Cohen's d = `r round(pwc4$estimate[1],digits = 2)`), supporting Hypothesis 2c.

@@

Since illusion data violated assumptions for parametric tests, an exploratory Friedman test was run and found a significant overall effect of condition with a moderate effect size ($\chi^2$(`r round(res.fried_illusion$df[1],digits = 2)`) = `r round(res.fried_illusion$statistic[1],digits = 2)`, *p* = `r format.pval(res.fried_illusion$p[1], digits = 1, eps = 0.001, nsmall = 3)`, Kendall’s W = `r round(effectsize_illusion$effsize[1],digits = 2)`) and post hoc Wilcoxon tests with Holm corrections found significantly greater combined illusion score in the Multisensory Stretching (MS) condition (M = `r round(illusionmsds$median[3],digits = 2)`, SD = `r round(illusionmsds$sd[3],digits = 2)`) compared to the Non-Illusion (NI; M = `r round(illusionmsds$median[1],digits = 2)`, SD = `r round(illusionmsds$sd[1],digits = 2)`, *z* = `r round(nppwc_illusion$statistic[2], digits = 2)`, *p.adj* = `r format.pval(nppwc_illusion$p.adj[2], digits = 1, eps = 0.001, nsmall = 3)`, r = `r round(nppwc_illusion$estimate[2],digits = 2)`), Non Illusion Tactile (NIT; M = `r round(illusionmsds$median[2],digits = 2)`, SD = `r round(illusionmsds$sd[2],digits = 2)`, *z* = `r round(nppwc_illusion$statistic[4], digits = 2)`, *p.adj* = `r format.pval(nppwc_illusion$p.adj[4], digits = 1, eps = 0.001, nsmall = 3)`, r = `r round(nppwc_illusion$estimate[4],digits = 2)`) and UV conditions (UV; M = `r round(illusionmsds$median[4],digits = 2)`, SD = `r round(illusionmsds$sd[4],digits = 2)`, *z* = `r round(nppwc_illusion$statistic[6], digits = 2)`, *p.adj* = `r format.pval(nppwc_illusion$p.adj[6], digits = 1, eps = 0.001, nsmall = 3)`, r = `r round(nppwc_illusion$estimate[6],digits = 2)`).

@@

This study sought to understand both subjective and neural responses to resizing illusions in participants without chronic pain, to provide not only a greater understanding of how bodily illusions affect cortical representations, but also a basis for investigating differences in cortical representations between participants with and without chronic pain conditions when using resizing illusions for analgesic treatment. Subjective data replicated previous findings of greater subjective illusory experience in multisensory compared to non-illusion conditions, showing that the addition of vibrotactile stimulation does not appear to impact subjective experience of resizing illusions using augmented reality, since these effects replicate ones found previously without vibrotactile stimulation. EEG data showed no significant effect of condition when assessing SSEP amplitudes across the electrodes of interest (F1 & FC1) at 26Hz. Exploratory non-parametric analyses of SSEP amplitudes showed a significant effect of condition however, with a decreased median amplitude in the multisensory and unimodal visual conditions compared to the non-illusion condition. However, these differences did not reach statistical significant with exploratory post hoc tests. These findings, therefore, demonstrate that within exploratory analyses illusory resizing can lead to reductions in SSEP amplitude, but this finding would need to be replicated as a confirmatory analysis and a larger sample size would likely be needed to detect any significant differences in post hoc comparisons due to the small effect sizes found.

>>>>

# install R packages
packagelist <- c('tidyverse','ggpubr','ez','ggplot2','jpeg','grid','plotrix','lavaan', 'cowplot', 'ggpubr','rstatix','reticulate','moments','tiff','rmarkdown','tinytex','osfr','FSA','compute.es')  
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

eegMSNIes <- tes(1.7,46,46)
eegUVNIes <- tes(1.15,46,46)
eegNINITes <- tes(2.02,46,46)

@@

nppwc_eeg1_es <- eegdata_2 %>%
  wilcox_effsize(Amplitude ~ Condition, paired = TRUE,comparisons = list(c("V1", "V3"), c("V1", "V4"), c("V1", "V2")))

@@

MSNIes <- tes(-5.67,46,46)
MSNITes <- tes(-5.61,46,46)

@@

nppwc_illusion_es <- subjecitve_data_illusion_1 %>%
  wilcox_effsize(Score ~ Condition, paired = TRUE)

@@

nppwc_disownership_es <- subjecitve_data_disownership_1 %>%
  wilcox_effsize(Score ~ Condition, paired = TRUE)

@@

Resizing illusions, delivered using augmented reality, resize a body part through either stretching or shrinking manipulations. These resizing illusions have been investigated in visuotactile, visual-only, and visuo-auditory presentations. However, the neural underpinnings of these resizing illusions remain undefined. This study sought to understand the neural mechanisms behind these illusions by using somatosensory steady state evoked potentials in addition to subjective self-report questionnaires, to enhance knowledge of what drives the subjective embodiment during resizing illusions. Since these Illusions have been shown to provide analgesic effects for individuals with chronic pain conditions, this study also aimed to provide an empirical basis for future investigations in chronic pain samples undergoing resizing illusions. Confirmatory analyses (N = 46) demonstrated significant differences in subjective experience between non-illusion and multisensory illusion conditions, and electroencephalography (EEG) data measuring somatosensory steady state evoked potential (SSEP) response across electrodes of interest (F1 & FC1) to 26Hz stimulation of the resized digit showed no significant effects of condition. However, further exploratory non-parametric SSEP analyses revealed a significant effect of condition, with reduced amplitudes in illusion conditions compared to non-illusion conditions, but no significant differences in exploratory post hoc tests. While confirmatory findings demonstrated no clear effect of resizing illusions on SSEP amplitudes for participants without chronic pain, exploratory findings could be interpreted as a potential “sharpening” of neural representations resulting from illusory stretching. These findings therefore provide a basis for investigations of comparable subjective and steady state illusion responses in a chronic pain population, who are thought to have more diffuse neural representations of their affected body parts.

@@

The augmented reality system used to deliver these resizing illusions presents real-time video capture of the hand, from the same position and perspective as if the hand were being viewed directly [@preston2011a]. This allows the experimenter to deliver tactile manipulations, such as gently pulling or pushing the hand/finger, whilst the participant views their hand/finger either stretching or shrinking in the augmented image. Newport, Pearce and Preston (2010) found strong embodiment using a synchronous multisensory visuotactile illusion, which was replicated in our pilot data using the same experimental set up as the current study. The pilot data showed, although not significant, a numerically greater illusory experience during synchronous visuotactile manipulations compared to asynchronous (mismatching visuotactile manipulation) control conditions (Appendix A) for illusory finger resizing. When comparing multisensory visuotactile resizing illusions to unimodal visual resizing illusions, our recent work [@hansford2023a] shows that multisensory illusions elicit significantly greater illusory experience compared to non-illusion and unimodal visual illusion conditions in healthy participants. We also showed, in exploratory analysis, that a subset of participants who experienced an illusion in the unimodal visual condition reported a stronger illusory experience in this condition than in an incongruent (mismatching visual and tactile inputs) control condition. Furthermore, we have demonstrated that a visuo-auditory presentation of the finger resizing illusion, using non-naturalistic auditory input, provides a stronger illusory experience than a visual only presentation, but this does not surpass the illusion strength given by a visuo-tactile illusion [@hansford2023b].

@@

## Preregistration

This study was preregistered as a stage 1 registered report which was given in-principal acceptance (IPA) by PCI-RR and be seen at the following OSF page: https://osf.io/pfksu/ . Due to the exploratory nature of the study and thereby the need for some slight methodological deviations, this study was withdrawn as a registered report during stage 2 revisions. All data and code to reproduce the analyses and manuscript for this study can be found at the following OSF page:  https://osf.io/yhz6j/.

## Deviations from Preregistration

-	During IPA there was an artefact from a previous round of revisions stating that if a participant needed either of the electrodes of interest removed due to noise, that the participant’s whole dataset would be removed. This should have stated that a participant’s dataset would only be removed if both electrodes of interest needed removal, otherwise analyses would be run using the remaining electrode of interest. The latter was the approach taken with SSEP data and analyses.

-	2.	It was preregistered that parametric analyses would be run to assess all hypotheses; however, the data did not meet all assumptions for parametric tests and therefore non-parametric tests were run instead, following standard statistical practice.

@@

Overall, based on the power analyses in section 2.7, a total sample size of 46 participants were tested. This sample size adheres to the higher end of sample size estimates (Hypothesis 2 (2.7.2) showing 46 participants needed for post hoc tests 2a – 2c).

@@

Ethical approval for this research was gained from the Department of Psychology, University of York (ethics application code 950), in line with the Declaration of Helsinki. All participants gave informed written consent prior to the start of any experimental set up, and participants were instructed that they could withdraw their participation at any time during or after completion of the experiment. 46 participants were tested, with the participant’s manipulated finger being randomly split between use of either the index or middle finger. However, 2 participant’s data needed removal (>50% of electrodes requiring removal), and therefore 2 additional participants were tested to account for this missing data, both using the index finger as the manipulated digit, resulting in a final sample size of 46 participants (37 Female, 8 Male, 1 Prefer not say; Mean age = 20.3 years, age range = 18.3 – 32.7 years; 32 White, 11 Asian or Asian British, 3 Mixed or Multiple Ethnic Groups; Sample population = students at the University of York). 23 participants were tested using their index finger, the other half using their middle finger.

### Sample inclusion / exclusion criteria:

@@

-	Less than 100% of the experiment completed by a participant, more than 50% of electrodes for a single participant requiring removal from EEG data, or if both electrodes F1 and FC1 (electrodes of interest) required removal. More information about data removal can be found in section 2.6 Preprocessing Steps.

@@

The experimenter was seated opposite the participant, the other side of the augmented reality machine and touched the digit during MS and NIT conditions by holding onto the distal interphalangeal joint and gently touching (NIT) or pulling (MS) the finger whilst the participant kept their hand in place. Conditions were delivered across 4 blocks, with each block consisting of 24 trials of the same experimental condition, totalling 96 trials over all 4 blocks. The ordering of the blocks was randomised for each participant to prevent ordering effects. The experiment was programmed in, and the conditions randomised using MATLAB R2017a and the experimenter was informed of whether to pull the finger or to touch the finger via an indicative box displayed on the screen out of the participant’s view. If the box was blue, this indicated a need to pull the finger, if it was white it indicated a need to touch the finger, if there was no box displayed then this indicated no tactile manipulation from the experimenter. The researcher used a button press to trigger the start of the manipulation, and started pulling the finger, when needed, synchronously within the 2.4 second manipulation phase. If the experimenter were to forget to pull the finger on a multisensory condition, or mistakenly pulled the finger in a control trial, then this would noted during the experiment, and that trial would be removed from analysis. Fortunately, no trials needed removal due to experimenter error. Vibrations were delivered to the participant’s finger in all conditions using a miniature electromagnetic solenoid stimulator (Dancer Design Tactor; diameter 1.8mm) emitting vibrations produced by sending amplified 26Hz sine wave sound files, with stimulus intensity controlled by an amplifier (Dancer Design TactAmp). The tactor was driven at 50% of the maximum (i.e. a peak input voltage of 3V) using a 26Hz sine-wave, and delivered a peak force of 0.18N. The electromagnetic solenoid stimulator was attached to the participant’s finger that was outstretched to receive the manipulations using clear medical tape, between the knuckle and the first finger joint, and gave continuous stimulation for the duration of each trial. Participants were encouraged to take a break between each of the blocks to stretch their hand. EEG was recorded throughout as a continuous recording with conditions denoted by numbered 8-bit digital at the start of each trial (USB-TTL Module, Black Box Toolkit Ltd.).

Finally, at the end of each block, the participant was asked to complete the subjective illusory experience questionnaire regarding a condition presented in a given block using a Samsung Galaxy Tab A6 tablet via a questionnaire on Qualtrics (Qualtrics, Provo, UT). The questionnaire consisted of six questions relating to the trials the participant had just experienced. Two statements related to illusory experience: “It felt like my finger was really stretching” / “It felt like the finger I saw was part of my body”, two related to disownership: “It felt like the finger I saw no longer belonged to me” / “It felt like the finger I saw was no longer part of my body”, and two were control questions: “It felt as if my finger had disappeared” / “It felt as if I might have had an extra finger” (all questions were directed towards the participants manipulated finger). Control questions were included to create an index for the illusion and disownership questions (more detail can be found in section 2.6 - Preprocessing steps), whilst disownership questions were included to assess if the potential experience from the illusions resulted from a disownership of the body part, or from subjective embodiment of the body part [@mccabe2011a]. A visual analogue scale from 0 – 100 was used for each statement, with 0 being strongly disagree, 50 being neutral and 100 being strongly agree.

@@

## Preprocessing Steps

@@

Regarding questionnaire data, scores for both illusion experience questions were combined to give median scores, along with both disownership questions and both control questions, resulting in 3 median scores per condition per participant. The median control scores were used to create an index of the illusion and disownership scores by subtracting the median control score from the median illusion and median disownership scores, in line with previous research doing similarly [@matsumiya2021a; @kilteni2017a; @kalckert2012a]. The normalised (control indexed) data were used for analyses, with a new scale from -100 to +100 with 100 indicating strongly agree, 50 indicating a neutral opinion, and scores below 0 indicating strongly disagree with the statements on the questionnaire. 50 was maintained as a neutral opinion so that the normalised data still adhered to the thresholds that the participants were presented with during the experiment.

All planned analyses can be seen within the stage 1 IPA report at the following OSF page: https://osf.io/pfksu/.

@@

@@

Positive control analyses of the subjective illusion data can be seen in @fig-illusion. A one-way ANOVA found a significant overall effect of condition with a large effect size (*F*(`r round(res.aov1$ANOVA$DFn[1])`,`r round(res.aov1$ANOVA$DFd[1])`), *p.adj* = `r format.pval(res.aov1$ANOVA$p[1], digits = 1, eps = 0.001, nsmall = 3)`, $ηp^2$ = `r res.aov1$ANOVA$pes[1]`). Post hoc t tests with Bonferroni corrections found significantly greater combined illusion score in the MS condition (M = `r round(illusionmsds$mean[3],digits = 2)`, SD = `r round(illusionmsds$sd[3],digits = 2)`) compared to the Non-Illusion (NI; M = `r round(illusionmsds$mean[1],digits = 2)`, SD = `r round(illusionmsds$sd[1],digits = 2)`, *t* = `r round(pwc1$statistic[2], digits = 2)`, *p.adj* = `r format.pval(pwc1$p.adj[2], digits = 1, eps = 0.001, nsmall = 3)`, Cohen's d = `r round(MSNIes$d[1], digits = 2)`) and Non Illusion Tactile (NIT; M = `r round(illusionmsds$mean[2],digits = 2)`, SD = `r round(illusionmsds$sd[2],digits = 2)`, *t* = `r round(pwc1$statistic[4], digit = 2)`, *p.adj* = `r format.pval(pwc1$p.adj[4], digits = 1, eps = 0.001, nsmall = 3)`, Cohen's d = `r round(MSNITes$d[1], digits = 2)`) conditions, thereby supporting hypotheses 1, 1a, and 1b and fulfilling the positive control checks.

@@

Analyses of SSEP data can be seen in @fig-eeg1. The left panel confirms the presence of a clear steady-state signal at 26Hz, which was strongest over the fronto-central electrodes. A one-way ANOVA found no significant effect of condition with a small effect size (*F*(`r round(res.aov4$ANOVA$DFn[1])`,`r round(res.aov4$ANOVA$DFd[1])`), *p* = `r format.pval(res.aov4$ANOVA$p[1], digits = 1, eps = 0.001, nsmall = 3)`, $ηp^2$ = `r res.aov4$ANOVA$pes[1]`), opposing Hypothesis 2. Post hoc t tests with Bonferroni corrections found no significant differences between SSEP amplitude when comparing the NI condition (M = `r round(eegdatamsds$mean[1],digits = 2)`, SD = `r round(eegdatamsds$sd[1],digits = 2)`) to the MS condition (M = `r round(eegdatamsds$mean[3],digits = 2)`, SD = `r round(eegdatamsds$sd[3],digits = 2)`, *t* = `r round(pwc4$statistic[2], digits = 2)`, *p.adj* = `r format.pval(pwc4$p.adj[2], digits = 1, eps = 0.001, nsmall = 3)`, Cohen's d = `r round(eegMSNIes$d[1],digits = 2)`), or UV condition (M = `r round(eegdatamsds$mean[4],digits = 2)`, SD = `r round(eegdatamsds$sd[4],digits = 2)`, *t* = `r round(pwc4$statistic[3], digits = 2)`, *p.adj* = `r format.pval(pwc4$p.adj[3], digits = 1, eps = 0.001, nsmall = 3)`, Cohen's d = `r round(eegUVNIes$d[1],digits = 2)`), meaning Hypotheses 2a and 2b were unsupported. There was no significant difference found when comparing the NI condition to the NIT condition (M = `r round(eegdatamsds$mean[2],digits = 2)`, SD = `r round(eegdatamsds$sd[2],digits = 2)`, *t* = `r round(pwc4$statistic[1], digits = 2)`, *p.adj* = `r format.pval(pwc4$p.adj[1], digits = 1, eps = 0.001, nsmall = 3)`, Cohen's d = `r round(eegNINITes$d[1],digits = 2)`), supporting Hypothesis 2c.

@@

Since illusion data violated assumptions for parametric tests, an exploratory Friedman test was run and found a significant overall effect of condition with a moderate effect size ($\chi^2$(`r round(res.fried_illusion$df[1],digits = 2)`) = `r round(res.fried_illusion$statistic[1],digits = 2)`, *p* = `r format.pval(res.fried_illusion$p[1], digits = 1, eps = 0.001, nsmall = 3)`, Kendall’s W = `r round(effectsize_illusion$effsize[1],digits = 2)`) and post hoc Wilcoxon tests with Holm corrections found significantly greater combined illusion score in the Multisensory Stretching (MS) condition (M = `r round(illusionmsds$median[3],digits = 2)`, SD = `r round(illusionmsds$sd[3],digits = 2)`) compared to the Non-Illusion (NI; M = `r round(illusionmsds$median[1],digits = 2)`, SD = `r round(illusionmsds$sd[1],digits = 2)`, *z* = `r round(nppwc_illusion$statistic[2], digits = 2)`, *p.adj* = `r format.pval(nppwc_illusion$p.adj[2], digits = 1, eps = 0.001, nsmall = 3)`, r = `r round(nppwc_illusion_es$effsize[2],digits = 2)`), Non Illusion Tactile (NIT; M = `r round(illusionmsds$median[2],digits = 2)`, SD = `r round(illusionmsds$sd[2],digits = 2)`, *z* = `r round(nppwc_illusion$statistic[4], digits = 2)`, *p.adj* = `r format.pval(nppwc_illusion$p.adj[4], digits = 1, eps = 0.001, nsmall = 3)`, r = `r round(nppwc_illusion_es$effsize[4],digits = 2)`) and UV conditions (UV; M = `r round(illusionmsds$median[4],digits = 2)`, SD = `r round(illusionmsds$sd[4],digits = 2)`, *z* = `r round(nppwc_illusion$statistic[6], digits = 2)`, *p.adj* = `r format.pval(nppwc_illusion$p.adj[6], digits = 1, eps = 0.001, nsmall = 3)`, r = `r round(nppwc_illusion_es$effsize[6],digits = 2)`).

@@

This study sought to understand both subjective and neural responses to resizing illusions in participants without chronic pain, to provide not only a greater understanding of how bodily illusions affect cortical representations, but also a basis for investigating differences in cortical representations between participants with and without chronic pain conditions when using resizing illusions for analgesic treatment. Subjective data replicated previous findings of greater subjective illusory experience in multisensory compared to non-illusion conditions, showing that the addition of vibrotactile stimulation does not appear to impact subjective experience of resizing illusions using augmented reality, since these effects replicate ones found previously without vibrotactile stimulation. Confirmatory EEG analyses showed no significant effect of condition when assessing SSEP amplitudes across the electrodes of interest (F1 & FC1) at 26Hz. Exploratory non-parametric analyses of SSEP amplitudes, however, showed a significant effect of condition with a decreased median amplitude in the multisensory and unimodal visual conditions compared to the non-illusion condition. However, these differences did not reach statistical significant with exploratory post hoc tests. These findings, therefore, suggest that illusory resizing can lead to reductions in SSEP amplitude, but replication within confirmatory analyses is needed.
